---
title: "PBC Commodity Index"
sidebar: false
---

[Switch to Russian](pbc-index.qmd)

**PBC Commodity Index** - а composite price index that measures price changes of Russian commodity export comparing to the level of January 2022.

```{r, warning = FALSE, echo = FALSE, message = FALSE}

library(openxlsx)
library(highcharter)
library(tidyr)
library(magrittr)

colors = c("#1a4afc", '#28055a', '#6d7aa5', '#82096c', '#d1e5ff', '#44bf78', '#f6514c',  "#FF9500", '#c3b0ef', '#d0d0d0', '#ffffff', '#000000' )

# прочитать и преобразовать файл Excel с данными

df  <- read.xlsx('xls/ЦЦИ Сырьевой индекс.xlsx', sheet = 1, detectDates = TRUE)

df_index = df[1,]
#df = df[,-1]
#df[,2:3] <- NULL 
names(df_index)[1] <- "index"

df_index <- df_index %>% pivot_longer(!index, names_to = "date", values_to = 'value')
df_index$date <- as.Date(df_index$date)

#df = df[df$index == 'PBC Commodity Index, 01.2022=100",]
df_index$index = 'PBC Commodity Index, Jan 2022 = 100'
df_index$value <- round(df_index$value,2)

# прочитать файл с курсом

# df2  <- read.xlsx('rub-usd.xlsx', sheet = 1, detectDates = TRUE)
# df2 <- df2[df2$date > min(df$date),]
df_usdrub = df[2,]
names(df_usdrub)[1] <- 'index'

df_usdrub <- df_usdrub %>% pivot_longer(!index, names_to = "date", values_to = 'value')
df_usdrub$date <- as.Date(df_usdrub$date)

df_usdrub$index = 'USD/RUB'
df_usdrub$value <- round(df_usdrub$value,2)

# Non-oil index

df_noil_index = df[4,]
names(df_noil_index)[1] <- 'index'

df_noil_index <- df_noil_index %>% pivot_longer(!index, names_to = "date", values_to = 'value')
df_noil_index$date <- as.Date(df_noil_index$date)

df_noil_index$index = 'Non-oil Export Price Index, Jan 2022 = 100'
df_noil_index$value <- round(df_noil_index$value,2)

# Fossil fuel index

df_fuel_index = df[6,]
names(df_fuel_index)[1] <- 'index'

df_fuel_index <- df_fuel_index %>% pivot_longer(!index, names_to = "date", values_to = 'value')
df_fuel_index$date <- as.Date(df_fuel_index$date)

df_fuel_index$index = 'Energy products Price Index, Jan 2022 = 100'
df_fuel_index$value <- round(df_fuel_index$value,2)

# Agro index

df_agro_index = df[7,]
names(df_agro_index)[1] <- 'index'

df_agro_index <- df_agro_index %>% pivot_longer(!index, names_to = "date", values_to = 'value')
df_agro_index$date <- as.Date(df_agro_index$date)

df_agro_index$index = 'Agricultural commodities Price Index, Jan 2022 = 100'
df_agro_index$value <- round(df_agro_index$value,2)

# Fertilizers index

df_fert_index = df[8,]
names(df_fert_index)[1] <- 'index'

df_fert_index <- df_fert_index %>% pivot_longer(!index, names_to = "date", values_to = 'value')
df_fert_index$date <- as.Date(df_fert_index$date)

df_fert_index$index = 'Fertilizers Price Index, Jan 2022 = 100'
df_fert_index$value <- round(df_fert_index$value,2)


# построить график highcharts 

index_plot <- highchart() %>% 
   hc_xAxis(type = "datetime") %>%
   hc_exporting(enabled = FALSE) %>% 
   hc_yAxis_multiples(
      list(title = list(text = "PBC Commodity Index, Jan 2022 = 100"), showFirstLabel = TRUE, max = 150, showLastLabel=TRUE,opposite = FALSE),
      list(title = list(text = "USD/RUB"), showLastLabel = FALSE, opposite = TRUE, reversed = TRUE)) %>%
   hc_add_series(df_index,type="line", hcaes(x = date, y = value, group=index), color = colors[1]) %>% 
   hc_add_series(df_noil_index,type="line", hcaes(x = date, y = value, group=index), color = colors[2], visible = TRUE) %>%
   hc_add_series(df_fuel_index,type="line", hcaes(x = date, y = value, group=index), color = colors[12], visible = FALSE) %>%
   hc_add_series(df_agro_index,type="line", hcaes(x = date, y = value, group=index), color = colors[3], visible = FALSE) %>%
   hc_add_series(df_fert_index,type="line", hcaes(x = date, y = value, group=index), color = colors[4], visible = FALSE) %>%
   hc_add_series(df_usdrub,type="line", hcaes(x = date, y = value, group=index), color = colors[8],  yAxis=1, visible = FALSE)

# поменять названия месяцев на русские. по умолчанию - на английском
#index_plot$x$conf_opts$lang$months = c("Январь"	,"Февраль",	"Март"	,"Апрель"	,"Май",	"Июнь", "Июль",	"Август",	"Сентябрь",	"Октябрь",	"Ноябрь",	"Декабрь")
#index_plot$x$conf_opts$lang$shortMonths = c("Янв", 	"Фев",	"Мар",	"Апр",	"Май",	"Июн",	"Июл",	"Авг",	"Сен",	"Окт",	"Ноя",	"Дек")

index_plot 

```

------------------------------------------------------------------------

**PBC Commodity Index change to the previous week: [0.0 п.]{.my-custom-class style="color:44bf78;"}** ([**-4.8 п.**]{.my-custom-class style="color:#f6514c;"} over the previous four weeks)

The biggest commodity price changes over the week: Crude Oil [**+0.8%**]{.my-custom-class style="color:#44bf78;"}, Natural Gaz [**-1.7%**]{.my-custom-class style="color:#f6514c;"}

Over the week the Index has been mainly affected by: Crude Oil [**-3.0 p.**]{.my-custom-class style="color:#f6514c;"}, Natural Gaz [**-1.2 p.**]{.my-custom-class style="color:#f6514c;"}

```{r, warning = FALSE, echo = FALSE, message = FALSE}
df_eng  <- read.xlsx('xls/ЦЦИ Сырьевой индекс.xlsx', sheet = 2, detectDates = TRUE)

df_factor <- df_eng[12:21,] %>% pivot_longer(-X1, names_to = "date", values_to = 'value')
df_factor$date <- as.Date(df_factor$date)
df_factor <- subset(df_factor, df_factor$date>'2024-02-01')

df_diff <- df_eng[11,] %>% pivot_longer(-X1, names_to = "date", values_to = 'value')
df_diff$date <- as.Date(df_diff$date)
df_diff <- subset(df_diff, df_diff$date>'2024-02-01')

factor_plot <- highchart() %>% 
   #hc_chart(type  = 'column') %>%
   hc_plotOptions(column = list(stacking = 'normal')) %>%
   hc_exporting(enabled = FALSE) %>% 
   hc_xAxis(type = 'datetime') %>%
   hc_yAxis(title = list(text = "Commodities impact in Index change, 4-weeks cumulative, points")) %>%
   hc_add_series(df_factor, hcaes(x = date, y = round(value,2), group = X1), type = 'column', pointWidth = 12, color = colors[1:10]) %>%
   hc_add_series(df_diff, hcaes(x = date, y = round(value,2), group = X1), type = 'line', marker = FALSE, color = colors[12])

# поменять названия месяцев на русские. по умолчанию - на английском
#factor_plot$x$conf_opts$lang$months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
#factor_plot$x$conf_opts$lang$shortMonths = c("Jan", "Feb",	"Mar",	"Apr",	"May",	"Jun",	"Jul",	"Aug",	"Sep",	"Oct",	"Nov",	"Dec")

factor_plot

```

------------------------------------------------------------------------

**Price differential (discount)** - the difference in price dynamics in Russian main export commodities compared to the prices of world benchmarks.

```{r, warning = FALSE, echo = FALSE, message = FALSE}

# World index

df_index_world = df[9,]
names(df_index_world)[1] <- 'index'

df_index_world <- df_index_world %>% pivot_longer(!index, names_to = "date", values_to = 'value')
df_index_world$date <- as.Date(df_index_world$date)

df_index_world$index = 'Commodity Index (World Prices), Jan 2022 = 100'
df_index_world$value <- round(df_index_world$value,2)

# Discount

df_discount = df[10,]
names(df_discount)[1] <- 'index'

df_discount <- df_discount %>% pivot_longer(!index, names_to = "date", values_to = 'value')
df_discount$date <- as.Date(df_discount$date)

df_discount$index = 'Discount, %'
df_discount$value <- round(df_discount$value,2)

discount_plot <- highchart() %>% 
   hc_xAxis(type = "datetime") %>%
   hc_exporting(enabled = FALSE) %>% 
   hc_yAxis_multiples(
      list(title = list(text = "PBC Commodity Index, Jan 2022 = 100"), showFirstLabel = TRUE, max = 150, showLastLabel=TRUE,opposite = FALSE),
      list(title = list(text = "Dicount, %"), showLastLabel = FALSE, opposite = TRUE, max = 50, min = 0)) %>%
   hc_add_series(df_index,type="line", hcaes(x = date, y = value, group=index), color = colors[1]) %>% 
   hc_add_series(df_index_world,type="line", hcaes(x = date, y = value, group=index), color = colors[2], visible = TRUE) %>%
   hc_add_series(df_discount,type="column", hcaes(x = date, y = value, group=index), color = colors[3],  yAxis=1, visible = TRUE)

# поменять названия месяцев на русские. по умолчанию - на английском
#discount_plot$x$conf_opts$lang$months = c("Январь"	,"Февраль",	"Март"	,"Апрель"	,"Май",	"Июнь", "Июль",	"Август",	"Сентябрь",	"Октябрь",	"Ноябрь",	"Декабрь")
#discount_plot$x$conf_opts$lang$shortMonths = c("Янв", 	"Фев",	"Мар",	"Апр",	"Май",	"Июн",	"Июл",	"Авг",	"Сен",	"Окт",	"Ноя",	"Дек")

discount_plot


```

------------------------------------------------------------------------

### [Monthly overview]{.my-custom-class style="color:#1a4afc;"}

April 2, 2025. PBC Commodity price index. **Commodity prices are pessimistic about global growth**

[Materials -\>](reports/2025-03 PBC Commodity Index_eng.pdf)


### [Methodology]{.my-custom-class style="color:#1a4afc;"}

[PBC Commodity index methodology -\>](pdf/PBC_Commodity_Index_Methodology.pdf)
